---
title: "DATA 607 - Assignment 5"
author: "Jim Mundy"
output:
  html_document:
    css:
    highlight: pygments
    theme: cerulean
    toc: true
    toc_float: true
  pdf_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r echo=TRUE, warning=FALSE, message=FALSE}
library(tidyverse)
library(kableExtra)
library(hrbrthemes)
library(reshape2)

```

## My Approach

My approach follows:

* Created csv file per assignment requirments.
* Used Readr to read the file into R
* Utilized Tidyverse tools to tidy and transform data.
* Presented data in table and graph formats.

## Read Data into R

```{r echo=TRUE, warning=FALSE, message=FALSE}

delay_data <- read_csv("https://raw.githubusercontent.com/MundyMSDS/DATA607/master/airline_delays.csv")

head(delay_data)
```

## Tidy the Data

```{r echo=TRUE, warning=FALSE, message=FALSE}


delay_data <- delay_data %>% 
  mutate(X1=if_else(is.na(X1),lag(X1),X1)) %>%
  drop_na() %>% 
  gather(`Los Angeles`, Phoenix, `San Deigo`, `San Francisco`, Seattle, key="City", value = "Count") %>% 
  spread("X2", "Count",  3:7) %>% 
  rename(Airline = X1, Delayed = delayed, `On Time` = `on time`) %>% 
  mutate(Total = Delayed + `On Time`) %>% 
  mutate(`On Time %` = `On Time` / Total) %>% 
  mutate(`Delay %` = Delayed / Total) %>%
  mutate(`Is Hub` = if_else(Airline=="Alaska" & City =="Seattle" |Airline=="AM West" & City =="Phoenix", "Yes", "No" )) %>% 
  arrange(Airline, Delayed)



```


## Analysis

```{r echo=TRUE, warning=FALSE, message=FALSE}
kable(delay_data, format = "markdown")

delay_data <- delay_data %>% 
  select(Airline, City, Delayed, `On Time`) %>%
  mutate(City = if_else(City == "Los Angeles", "LA", City)) %>% 
  mutate(City = if_else(City == "San Francisco", "SF", City)) %>% 
  mutate(City = if_else(City == "San Deigo", "SD", City)) %>%
  mutate(City = if_else(City == "Phoenix", "PHX", City)) %>%
  mutate(City = if_else(City == "Seattle", "STL", City)) %>%
  melt()

ggplot(delay_data, aes(x = City, y = value, fill = variable)) +
   geom_bar(stat = 'identity') +
   geom_text(size = 3, aes(label = value),  position = position_stack(vjust = 0.5)) +
   labs(y = "flights") +
   labs(x="", y="Count",
       title="Flight Delays",
       subtitle="Alaska Air vs. America West",
       caption="Brought to you by the Tidyverse") + 
   theme_ipsum(grid="Y") +
   facet_grid(~ Airline)

  

```



## Findings

* More flights mean more delays - Hubs (Seattle:Alaska, Phoenix:Am West) had highest delays for each airline.

* Alaska Airlines has a better track record for delays versus AM West.  

* Alaska Air's highest on time percentage is at American West's hub (Phoenix)

* Delays are positively correlated to precipitation. 



