---
title: "DATA 607 - Assignment 3"
author: "Jim Mundy"
output:
  html_document:
    css: ./lab.css
    highlight: pygments
    theme: cerulean
    toc: true
    toc_float: true
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```


```{r echo=TRUE, warning=FALSE, message=FALSE}
library(tidyverse)
library(dbplyr)
library(kableExtra)


data607RegexChecker <- function(regex_strings,regex_pattern){
  
m <- str_extract_all(regex_strings,pattern=regex_pattern, simplify = TRUE)
m <- rename(as_tibble(m), 'match' = "V1")

m = m %>%
 mutate(String = regex_strings) %>%
 mutate(Match = ifelse(match=="",'No Match',match)) %>% 
 select(String, Match)

 kable(m, format = "markdown")  
  
}


```

###3. Create Strings. 

Use the following string to complete the exercises:

```{r}
names <-"555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"
```



```{r echo=TRUE, warning=FALSE, message=FALSE}
names <- str_extract_all(names,'[[a-zA-Z],. ]{2,}')
names <- as_tibble(names, .name_repair="universal")
names <- rename(names, 'name_strings'= '...1')

names <- names %>%
  mutate(Number = row_number() ) %>%
  mutate(title = str_extract(name_strings,"^[a-zA-Z]{2,4}\\.")) %>% 
  mutate(name_strings = str_replace_all(name_strings, "^[a-zA-Z]{2,4}\\.\\s",'' )) %>% 
  mutate(first_name = str_trim(str_extract(name_strings,'^[a-zA-Z]+\\s?\\,?'))) %>% 
  mutate(last_name = str_trim(str_extract(name_strings,'(\\s.*)$'))) %>% 
  mutate(last_name = if_else(str_detect(first_name,'\\,'),first_name,last_name)) %>% 
  mutate(first_name = if_else(str_detect(first_name,'\\,'),str_trim(str_extract(name_strings,'(\\s.*)$')),first_name)) %>% 
  mutate(last_name = str_replace(last_name,'\\,','')) %>%
  mutate(first_name_last_name = if_else(is.na(title),str_c(first_name, last_name, sep=' '),str_c(title, first_name, last_name, sep=' ')))
```

1. Use the tools of this chapter to rearrange the vector so that all the elements conform to the standard first_name last_name

```{r}
first_last <- names %>%
  select(first_name_last_name) 
  kable(first_last, format = "markdown")


```


2. Construct a logical vector indicating whether a character has a title (i.e. Rev., and Dr.)

```{r}
title_vector <- names %>%
  mutate(has_title = str_detect(first_name_last_name, '^[a-zA-Z]{2,4}\\.')) %>% 
  select(first_name_last_name, has_title)

kable(title_vector, format = "markdown")


```

3. Construct a logical vector indicating whether a character has a second name.

```{r}
second_name_vector <- names %>% 
  mutate(has_second_name = str_detect(first_name_last_name, '^[A-Z\\.]{2}\\s[a-zA-Z]+\\s[a-zA-z]+$')) %>% 
  select(first_name_last_name, has_second_name)

kable(second_name_vector, format = "markdown")

```


###4. Types of Strings

Describe the types of strings that conform to the following regular expressions and construct an example that is matched by the regular expression. 


1. Regular Expression Pattern:  [0-9]+\\$

Description: One to unlimited digits followed by a single dollar sign ($).

For example: 
```{r}

(data607RegexChecker(c("0$", "000r00?", "123456$", "1111111111$", "go_yankees!"),
                     "[0-9]+\\$" ))

```


2. Regular Expression Pattern: \\b[a-z]{1,4}\\b

Description: This will match lower case character strings of length 1 to 4.

For example: 
```{r}

(data607RegexChecker(c("a", "abcd", "abcD", "go", "yank"),
                     "\\b[a-z]{1,4}\\b" ))

```


3. Regular Expression Pattern: .*?\\.txt$

Description: Matches any character except line termination 0 to unlimited times, followed by a period and finally followed by the characters txt. This appears to be a DOS statement that looking for all files that end in .txt.

For example: 

```{r}

(data607RegexChecker(c("resume.txt", "memo.txt", "spreadsheet.xls", "go.txt", "file.pdf"),
                     ".*?\\.txt$" ))

```



4. Regular Expression Pattern: \\d{2}/\\d{2}/\\d{4}

Description: Two digits followed by a back slash followed by Two digits followed by a back slash followed by four digits. In other word this could represent a tranditional date format. 

For example: 

```{r}

(data607RegexChecker(c("04/07/1963", "05/19/1998", "09/11/2001", "8/7/1980", "09/15/2019"),
                     "\\d{2}/\\d{2}/\\d{4}" ))

```



5. Regular Expression Pattern: <(.+?)>.+?</\\1>

Description: This matches html data.Intially it matches the beginning tag, any 1 to unlimited character(s) inside the first brackets (<>), next it matches any 1 to unlimited characters (between the brackets) and finally it matches the ending brackets and uses the result from the 1st capture group (\\1) in the end bracket( </\\1> ), thus eliminating the need to match something that has already been matched.

```{r}

m <- str_extract_all(c("<H1> This is a header </H1>", "<p> This is a paragraph </p>","<date> 12/13/234</date>", "<ul>unordered list</ul>", "Date"),"<(.+?)>.+?</\\1>", simplify = TRUE)

m
```


6. Bonus:  Shinny R-flavored Regex Tester

Here is a link to a shiny regex tester that allows you to use R-flavored regular expressions. I used it to test my expressions.

https://adamspannbauer.github.io/2018/01/16/r-regex-tester-shiny-app/




###9. Secret Message

The following code hides a secret message. Crack it with R and regular expressions.  Hint: Some of the characters are more revealing than others! The code snippet is also available in the materials at www.r-datacollection.com

```{r}
code = "clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0TanwoUwisdij7Lj8pf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigOd6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr"
```

There appears to be more lowercase characters than uppercase, so I will strip all the lower case characters and see what that yields:

```{r}
secret_message <- str_replace_all(code,'[a-z\\s\\d]+','')
secret_message <- str_replace_all(secret_message,'[\\.]',' ')

```

**THE SECRET MESSAGE IS:** `r secret_message`.
