---
title: "DATA 607 - Assignment 4"
author: "Jim Mundy"
output:
  html_document:
    css: ./lab.css
    highlight: pygments
    theme: cerulean
    toc: true
    toc_float: true
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r echo=TRUE, warning=FALSE, message=FALSE}
library(tidyverse)
library(kableExtra)

```

## My Approach

My approach for this project follow:

* Import data using readr (used fixed length file)
* Use the Tidyverse to wrangle data into the desired form
* Write a CSV file using readr
* Import csv file using Shiny App 

Note: Inspiration and code for shiny app come from https://www.youtube.com/watch?v=HPZSunrSo5M

## Import Data

```{r echo=TRUE, warning=FALSE, message=FALSE}

columns <- c("player","c2","c3","c4","c5","c6","c7","c8","c9","c10")
delo <- read_fwf("https://raw.githubusercontent.com/MundyMSDS/DATA607/master/elodataset.txt", fwf_widths(c(6, 32, 7,7,6,6,6,6,6,6), columns),skip=4, comment="-")
glimpse(delo)
```

## Wrangle Data

```{r echo=TRUE, warning=FALSE, message=FALSE}

d <- delo %>% 
  mutate(c2= str_trim(str_sub(c2,2,29),side="both")) %>% 
  mutate(c3 = str_trim(str_sub(c3,2,6),side="both")) %>%
  mutate(c4 = str_trim(str_sub(c4,2,6),side="both")) %>%
  mutate(c5 = str_trim(str_sub(c5,2,6),side="both")) %>%
  mutate(c6 = str_trim(str_sub(c6,2,6),side="both")) %>%
  mutate(c7 = str_trim(str_sub(c7,2,6),side="both")) %>%
  mutate(c8 = str_trim(str_sub(c8,2,6),side="both")) %>%
  mutate(c9 = str_trim(str_sub(c9,2,6),side="both")) %>%
  mutate(c10 = str_trim(str_sub(c10,2,6),side="both")) %>% 
  mutate(c11 = if_else(str_detect(player,"\\d"),lead(player),"")) %>% 
  mutate(c12 = if_else(str_detect(player,"\\d"),lead(c2),"")) %>% 
  filter(str_detect(player,"\\d")) %>% 
  mutate(c4=as.integer(str_replace_all(c4,"[WLDHU]\\s",""))) %>%
  mutate(c5=as.integer(str_replace_all(c5,"[WLDHU]\\s",""))) %>%
  mutate(c6=as.integer(str_replace_all(c6,"[WLDHU]\\s",""))) %>%
  mutate(c7=as.integer(str_replace_all(c7,"[WLDHU]\\s",""))) %>%
  mutate(c8=as.integer(str_replace_all(c8,"[WLDHU]\\s",""))) %>%
  mutate(c9=as.integer(str_replace_all(c9,"[WLDHU]\\s",""))) %>%
  mutate(c10=as.integer(str_replace_all(c10,"[WLDHU]\\s",""))) %>% 
  mutate(c12 = str_replace_all(c12,"\\d{8}\\s\\/\\sR\\:\\s","")) %>% 
  separate(c12, into=c("c13", "c14"), sep = "->", remove = FALSE) %>% 
  mutate(c13 = str_extract(c13,"\\d{3,4}"))

d2 <- d  %>% 
  select(c2,c11,c3, c13,c4:c10) %>% 
  gather(c4:c10, key="match", value="player") %>% 
  arrange(c2) %>%
  mutate(player = as.character(player)) %>% 
  select(-match)
  

d3 <- d %>% 
  select (player, c13)
 
result <- left_join(d2, d3, by ="player") %>% 
  group_by(c2, c11,c3, c13.x) %>%  
  summarize(score = mean(as.integer(c13.y), na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(score = as.integer(score)) %>% 
  rename(Name=c2, State=c11, Ttl_Points =c3, Pre_Rating = c13.x, Avg_Opp_Rating = score) %>% 
  arrange(desc(Avg_Opp_Rating))

```

## Elo Data Output

```{r echo=TRUE, warning=FALSE, message=FALSE}

  kable(result, format = "markdown")

```


## Write Data To CSV File

```{r echo=TRUE, warning=FALSE, message=FALSE}
  
  result <- format_csv(result)
  setwd(file.path("C:","Users", "mutue", "OneDrive", "Documents", "Data607"))
  write_file(result,"elo_data.csv")
  
```


## Shiny App


Click this link to access the Shiny App.  Use the App to access your ELO CSV file or other csv file. 

Link to App:  https://mundymsds.shinyapps.io/EloApp/




