---
title: "Data 607 - Assignment 1"
author: "James Mundy"
date: "September 1, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Get The Data

Use getURL function to get the data from my github repository. The assign the data to x. 

```{r echo=TRUE, results='hide', warning=FALSE, message=FALSE}
library(RCurl)

```

```{r echo=TRUE, results='hide', warning=FALSE, message=FALSE}
x <- getURL("https://raw.githubusercontent.com/MundyMSDS/DATA607/master/mushroom.csv")
```

## Import The Data

Will utlized the tidyverse library and related tools to complete this assignment.

```{r echo=TRUE, results='hide', warning=FALSE, message=FALSE}
library (tidyverse)
```


```{r echo=TRUE, results='hide', warning=FALSE, message=FALSE}
mushrooms <-read_csv(x, col_names= c("isEdible",
                                     "cap_shape",
                                     "cap_surface",
                                     "cap_color",
                                     "bruises",
                                     "odor",
                                     "gill_attachment",
                                     "gill_spacing",
                                     "gill_size",
                                     "gill_color",
                                     "stalk_shape",
                                     "stalk_root",
                                     "stalk_surface_above_ring",
                                     "stalk_surface_below_ring",
                                     "stalk_color_above_ring",
                                     "stalk_color_below_ring",
                                     "veil_type",
                                     "veil_color",
                                     "ring_number",
                                     "ring_type",
                                     "spore_print_color",
                                     "population",
                                     "habitat")
)

```

```{r}
glimpse(mushrooms)
```

## Tidy The Data

We will define each variable as a factor and then apply the required data transformations.

```{r}
mushrooms <- mushrooms %>% map_df(function(.x) as.factor(.x))

levels(mushrooms$isEdible) <- c("edible", "poisonous")
levels(mushrooms$cap_shape) <- c("bell", "conical", "flat", "knobbed", "sunken", "convex")
levels(mushrooms$cap_color) <- c("buff", "cinnamon", "red", "gray", "brown", "pink", 
                                 "green", "purple", "white", "yellow")
levels(mushrooms$cap_surface) <- c("fibrous", "grooves", "scaly", "smooth")
levels(mushrooms$bruises) <- c("no", "yes")
levels(mushrooms$odor) <- c("almond", "creosote", "foul", "anise", "musty", "none", "pungent", "spicy", "fishy")
levels(mushrooms$gill_attachment) <- c("attached", "free")
levels(mushrooms$gill_spacing) <- c("close", "crowded")
levels(mushrooms$gill_size) <- c("broad", "narrow")
levels(mushrooms$gill_color) <- c("buff", "red", "gray", "chocolate", "black", "brown", "orange", 
                                  "pink", "green", "purple", "white", "yellow")
levels(mushrooms$stalk_shape) <- c("enlarging", "tapering")
levels(mushrooms$stalk_root) <- c("missing", "bulbous", "club", "equal", "rooted")
levels(mushrooms$stalk_surface_above_ring) <- c("fibrous", "silky", "smooth", "scaly")
levels(mushrooms$stalk_surface_below_ring) <- c("fibrous", "silky", "smooth", "scaly")
levels(mushrooms$stalk_color_above_ring) <- c("buff", "cinnamon", "red", "gray", "brown", "pink", 
                                              "green", "purple", "white", "yellow")
levels(mushrooms$stalk_color_below_ring) <- c("buff", "cinnamon", "red", "gray", "brown", "pink", 
                                              "green", "purple", "white", "yellow")
levels(mushrooms$veil_type) <- "partial"
levels(mushrooms$veil_color) <- c("brown", "orange", "white", "yellow")
levels(mushrooms$ring_number) <- c("none", "one", "two")
levels(mushrooms$ring_type) <- c("evanescent", "flaring", "large", "none", "pendant")
levels(mushrooms$spore_print_color) <- c("buff", "chocolate", "black", "brown", "orange", 
                                         "green", "purple", "white", "yellow")
levels(mushrooms$population) <- c("abundant", "clustered", "numerous", "scattered", "several", "solitary")
levels(mushrooms$habitat) <- c("wood", "grasses", "leaves", "meadows", "paths", "urban", "waste")
```

Here are the transformed facators

```{r}
str(mushrooms)
```

Remove factor (veil_type) with one level.

```{r}
mushrooms <- mushrooms %>% select(- veil_type)
```

Check for missing data / values

```{r}
map_dbl(mushrooms, function(.x) {sum(is.na(.x))})
```

Remove gill_attachment because it has missing values

```{r}
mushrooms <- mushrooms %>% select(- gill_attachment)
```

Select isEdible and four or five other columns.

```{r}
mushrooms <- mushrooms %>% select(isEdible, odor, cap_surface, cap_color, gill_color, cap_shape)
```

Show summary of the mushrooms dataframe (tibble)

```{r}
summary(mushrooms)
```



## Explore The Selected Data

Use ggplot2 to visulize the data to help determine the interesting variables. 

```{r}
ggplot(mushrooms, aes(x = cap_surface, y = cap_color, col = isEdible)) + 
  geom_jitter(alpha = 0.5) + 
  scale_color_manual(breaks = c("edible", "poisonous"), 
                     values = c("green", "red"))
```

Fibrous mushrooms appear the safest to eat.

```{r}
ggplot(mushrooms, aes(x = cap_shape, y = cap_color, col = isEdible)) + 
  geom_jitter(alpha = 0.5) + 
  scale_color_manual(breaks = c("edible", "poisonous"), 
                     values = c("green", "red"))
```

Bell shaped mushrooms seem like the safe bet. 

```{r}
ggplot(mushrooms, aes(x = gill_color, y = cap_color, col = isEdible)) + 
  geom_jitter(alpha = 0.5) + 
  scale_color_manual(breaks = c("edible", "poisonous"), 
                     values = c("green", "red"))
```

```{r}
ggplot(mushrooms, aes(x = isEdible, y = odor, col = isEdible)) + 
  geom_jitter(alpha = 0.5) + 
  scale_color_manual(breaks = c("edible", "poisonous"), 
                     values = c("green", "red"))
```



'''
***Inspiration***
1. R for Data Science
2. Machine Learning with R
3. ggplot2 (the book)
'''
