---
title: "DATA 607 - Project 2a - Horse Racing Data"
author: "Jim Mundy"
output:
  html_document:
    css: ./lab.css
    highlight: pygments
    theme: cerulean
    toc: true
    toc_float: true
  pdf_document: default
---

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
library(RCurl)
library(tidyr)
library(kableExtra)
knitr::opts_chunk$set(echo = TRUE)


```

## The Data

#### The data is horse racing past performance data.  This is a challenging data set:

* __Data is stacked___ -The data set includes several types of data and is stacked within the file. For example, there are records for Starters, Past Performances, Race, Exoctic Wagers, Information, and Workouts. Each record type is designated by a record(S, P, I, E, W).  The data is also relatively wide with more than 40 unnamed columns.
* __Missing Data__ - Some horses in field are international so there is incomplete data for these horse
* __Data Type Conversiion Required__ -  Initially all of the data is character, therefore some data will need to be converted to other types (data, integer, etc.)
* __New Variables required__ - additional feilds will be required so mutate will be employed to create these fields
* __One to Many Data__ - There is a natural one to many relationship to this data so I will attempt to create a list column to reflect this

## My Game Plan

#### My game plan for this data set follows:

* __Import Data__ - Use readr to import the data
* __Get To A Working Data Set__  - Remove record type and combine record type to get to a good starting point
* __Create some helper functions to be used in the Wrangling / Transform step
* __Wrangling / Transform Data__ Create headers, variables, select and convert data  
* __Remove Missing__ Remove the international horses with imcomplete data
* __Create List Column__ There is a natural 1 to Many (One horse has many past performances) to this data I will try to capture this in a tidy way
* __Analyze__ I will leverage the list column to calculate the desired metrics to answer the questions.

## My Questions

#### 1. What horses have the top 10 most wins?
#### 2. What horses have the top 10 win percentage?
#### 3. What horses have the top 10 in the money percentage?


## Transform Data

#### The steps to transform the dataset are set forth below:

1.  __Import Data__

#### Utilize readr and and its col_names and col_types parameter to import the file. Column C1 captures the stacked nature of the file. The letter designation indicates the type of record (S,R, I, W, etc.)

```{r echo=TRUE, warning=FALSE, message=FALSE}

p <- getURL("https://raw.githubusercontent.com/MundyMSDS/DATA607/master/pps.csv")

past_performances <- read_csv(p, col_names = c("c1","c2","c3","c4","c5","c6","c7","c8","c9","c10","c11","c12","c13","c14","c15","c16",
      
                                                                             "c17","c18","c19","c20","c21","c22","c23","c24","c25","c26","c27","c28","c29","c30","c31","c32","c33","c34","c35","c36","c37","c38","c39","c40","c41","c42","c43","c44","c45","c46"),
                   col_types = ("cccccccccccccccccccccccccccccccccccccccccccccc"))

past_performances

```


2.  __Get To A Good Starting Point__

#### Now that the data is imported, I will get to a good starting point by filtering to the records that I will utilize (P and I), ensure a Horse name exists on each row and adding a record_id to preserve the order of the file and provide a good point of reference.  

```{r echo=TRUE, warning=FALSE, message=FALSE}

past_performances <- past_performances %>%
  mutate(record_id = row_number() ) %>% 
  filter(c1 == 'P' | c1 == 'I') %>% 
  mutate(info= ifelse(c1=='I',c3,'')) %>%
  mutate(horse = ifelse(c1=='I', lead(c3,1),c3)) %>%
  mutate(horse = ifelse(c1=='I', lead(horse,1),c3)) %>%
  mutate(yr = ifelse (c1 == 'I' & str_sub(c2,1,1)=='2',c2,NA)) %>%
  mutate(record_id = row_number()) %>%
  separate(c8, into=c("race_cls", "race_price"), sep = ";", remove = FALSE) %>% 
  group_by(horse) %>% mutate(sub_record_id = seq(n()))

pp <- past_performances %>% 
  select(record_id, horse)
  

head(pp) %>%
  kable() %>%
  kable_styling()  

```

3.  __Create Helper Functions__

#### Here I create two helper functions that will use in conjunction with mutate to create two new variables past_performance_code and furlongs (integer),  Past performance code is a code that will uniquely identify a past performance line and furlongs is the distance of the race in furlongs ( a furlong is 1/8 of mile.)

```{r echo=TRUE, warning=FALSE, message=FALSE}

get_pp_code <- function(dt_str, track, race, finish) {
  d <- str_trunc(dt_str,2, side = "left", ellipsis = "")
  m <- str_sub(dt_str,5,2)
  m <- month(ymd(dt_str),label=TRUE, abbr=TRUE)
  yr <- str_trunc(dt_str,4, side = "right", ellipsis = "")
  yr <- str_trunc(yr,2, side = "left", ellipsis = "")
  r <- str_pad(race, 2, "left", "0")
  p <- str_pad(finish,2, "left", "0")
  s <- "_"
  x <- str_c(d,m,yr,s,r,track,p, sep="")
}


get_furlongs <- function(d) {
  case_when(
    str_detect(d,'3 Furlongs') ~ 300,
    str_detect(d,'3 1/2 Furlongs') ~ 350, 
    str_detect(d,'3 3/4 Furlongs') ~ 375,
    str_detect(d,'4 Furlongs') ~ 400,
    str_detect(d,'4 1/4 Furlongs') ~ 425,
    str_detect(d,'4 1/2 Furlongs') ~ 450,
    str_detect(d,'5 Furlongs') ~ 500,
    str_detect(d,'About 5 Furlongs') ~ 500,
    str_detect(d,'5 1/4 Furlongs') ~ 525,
    str_detect(d,'5 Furlongs 80 Yards') ~ 536,
    str_detect(d,'5 1/2 Furlongs') ~ 550,
    str_detect(d,'About 5 1/2 Furlongs') ~ 550,
    str_detect(d,'6 Furlongs') ~ 600,
    str_detect(d,'About 6 Furlongs') ~ 600,
    str_detect(d,'6 1/2 Furlongs') ~ 650,
    str_detect(d,'About 6 1/2 Furlongs') ~ 650,
    str_detect(d,'7 Furlongs') ~ 700,
    str_detect(d,'About 7 Furlongs') ~ 700,
    str_detect(d,'7 1/2 Furlongs') ~ 750,
    str_detect(d,'1 Mile') ~ 800,
    str_detect(d,'About 1 Mile') ~ 800,
    str_detect(d,'1 Mile 40 Yards') ~ 818,
    str_detect(d,'About 1 Mile 40 Yards') ~ 818,
    str_detect(d,'1 Mile 70 Yards') ~ 832,
    str_detect(d,'About 1 Mile 70 Yards') ~ 832,
    str_detect(d,'1 1/16 Miles') ~ 850,
    str_detect(d,'About 1 1/16 Miles') ~ 850,
    str_detect(d,'1 1/8 Miles') ~ 900,
    str_detect(d,'1 3/16 Miles') ~ 950,
    str_detect(d,'1 1/4 Miles') ~ 1000,
    str_detect(d,'1 5/16 Miles') ~ 1040,
    str_detect(d,'1 5/16 Miles') ~ 1050,
    str_detect(d,'1 3/8 Miles') ~ 1100,
    str_detect(d,'1 7/16 Miles') ~ 1150,
    str_detect(d,'1 1/2 Miles') ~ 1200,
    str_detect(d,'1 9/16 Miles') ~ 1250,
    str_detect(d,'1 5/8 Miles') ~ 1300,
    str_detect(d,'1 11/16 Miles') ~ 1350,
    str_detect(d,'1 3/4 Miles') ~ 1400,
    str_detect(d,'1 13/16 Miles') ~ 1450,
    str_detect(d,'1 7/8 Miles') ~ 1500,
    str_detect(d,'1 15/16 Miles') ~ 1550,
    str_detect(d,'2 Miles') ~ 1600,
    TRUE ~ 0
  )
}


```



4.  __Data Wrangling / Transform__

#### In this step I will continue to build out the tidy data set by renaming columns (trk_cde, jockey, off_odds, dist_str), creating new variables with mutate (distance, pp_code) and converting variable types (post, finish and field to integer)

```{r echo=TRUE, warning=FALSE, message=FALSE}


past_performances <- past_performances %>% 
  group_by(horse) %>% mutate(sub_record_id = seq(n())) %>% 
  mutate(yr = ifelse(sub_record_id == 1 & is.na(yr), toString(year(now())),yr)) %>%
  tidyr::fill(yr) %>% 
  ungroup(horse) %>% 
  arrange(record_id) %>%
  mutate(mth_day = c5) %>% 
  separate(c5, into=c("r_mth", "r_day"), sep = " ", remove = FALSE) %>%
  mutate(r_mth = match(r_mth, month.abb)) %>%
  mutate(r_mth = str_pad(r_mth,2,"left", "0")) %>% 
  mutate(date_str = str_c(yr,r_mth,r_day,sep='')) %>%
  arrange(record_id) %>%
  
  #renaming columns
  rename(trk_cde = c2) %>%
  rename(race_nbr = c7) %>%
  rename(jockey = c4) %>%
  rename(off_odds = c17) %>%
  rename(dist_str = c10) %>%
  
  #Using helper functions to create distance variable
  mutate(distance = get_furlongs(dist_str)) %>% 
  
  #renaming columns
  rename(surface = c11) %>%
  rename(race_txt = c14) %>%
  rename(surface_cond = c12) %>%
  rename(wt = c15) %>%
  rename(eq_med = c16) %>%
  rename(post = c18) %>%
  rename(field = c19) %>%
  rename(finish = c28) %>%
  
  #using helper function to create pp_cde column
  mutate(pp_cde = get_pp_code(date_str, trk_cde, race_nbr,finish)) %>%
  
  #convering serveral variables from character to integer
  mutate(post = as.integer(post)) %>%
  mutate(finish = as.integer(finish)) %>%
  mutate(field = as.integer(field)) %>%
  ungroup()

pp <- past_performances %>% 
  select(record_id, horse, pp_cde, race_cls, jockey, off_odds, distance, surface, surface_cond, post, field, finish)
  

head(pp) %>%
  kable() %>%
  kable_styling()  

```

5.  __Remove Row With Missing Data__

#### A subset of the data includes international horses. These horses have incomplete data sets and will be removed from the data.

```{r echo=TRUE, warning=FALSE, message=FALSE}

pp <- pp %>% 
  filter(!is.na(pp_cde))

pp <- pp %>% 
  select(horse, post, field, finish)
  
#select(horse, pp_cde, race_cls, jockey, off_odds, distance, surface, surface_cond, post, field, finish)

head(pp) %>%
  kable() %>%
  kable_styling()

```

6.  __Create A List Column__

#### Creating a list colunn will essentially enable me to create a column (data) that is a comprised of past performances for each horse, the list will have 1 row for each past performance (by horse - ie different numbers of rows for each horse) and a column for eqch variable. This can

```{r echo=TRUE, warning=FALSE, message=FALSE}

pp <-  pp %>% 
  #creating a list column
  group_by(horse) %>% 
  nest() %>% 
  
  #add new variables that will enable me to answer questions
  mutate(starts=map_int(data,nrow)) %>% 
  mutate(win =map_dbl(data, ~sum(.$finish==1, na.rm=TRUE))) %>%
  mutate(place =map_dbl(data, ~sum(.$finish<=2, na.rm=TRUE))) %>%
  mutate(show =map_dbl(data, ~sum(.$finish<=3, na.rm=TRUE))) %>%
  mutate(avg_finish =map_dbl(data, ~mean(.$finish, na.rm=TRUE))) %>%
  mutate(win_pct = win/starts) %>%
  mutate(itm_pct = show/starts) %>% 
  arrange(desc(win))

```


#### Now I should have a tibble with a data list column of past performances for each horse and summary stats of the past performances for each horse

```{r echo=TRUE, warning=FALSE, message=FALSE}
pp

```

## Answers To Questions

#### 1. What horses have the top 10 most wins?

```{r echo=TRUE, warning=FALSE, message=FALSE}

select(pp, horse, starts, win, place, show, win_pct, itm_pct, avg_finish) %>% 
head(10L,pp) %>%
  kable() %>%
  kable_styling()

```


#### 2. What horses have the top win percentage?

```{r echo=TRUE, warning=FALSE, message=FALSE}
select(pp, horse, starts, win, place, show, win_pct, itm_pct, avg_finish) %>% 
arrange(desc(win_pct)) %>% 
  head(10L,pp) %>%
  kable() %>%
  kable_styling()


```


#### 3. What horses have the top in the money percentage?

```{r echo=TRUE, warning=FALSE, message=FALSE}
select(pp, horse, starts, win, place, show, win_pct, itm_pct, avg_finish) %>% 
arrange(desc(itm_pct)) %>% 
  head(10L,pp) %>%
  kable() %>%
  kable_styling()


```