---
title: "DATA 607 - Project 2c - Census Data Income vs Educational Attainment 2018"
author: "Jim Mundy"
output:
  html_document:
    css: ./lab.css
    highlight: pygments
    theme: cerulean
    toc: true
    toc_float: true
  pdf_document: default
---

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
library(modelr)
library(RCurl)
library(tidyr)
library(kableExtra)
library(plotly)
knitr::opts_chunk$set(echo = TRUE)


```

## The Data

#### This is a relatively clean data set. Potential area for tidying include:

* __Some additional binning__ - The data set is comprised of three distinct, yet related, data sets.
* __Merged Header__ - The file has a merged header and will require some clean-up
* __Filter__ - Filtering rows that are not needed for analysis
* __Missing and Incomplete Data__ - There is a need to deal with incomplete and/or missing data
* __Skip Rows__ - Willl need to skip rows during data import due to blank space and a fat header.

## My Game Plan

#### My game plan for this data set follows:

* Import the data 
* Tidy the data
* Use ggplot to determine correlations and insights

## My Question

#### Indentify correlation between income and educational attainment.

## Transform Data

#### The steps to transform the dataset are set forth below:

(1)  __Import Data__

#### Utilize readr with skip parameter of 11 to import the data.  

```{r echo=TRUE, warning=FALSE, message=FALSE}

inc <- getURL("https://raw.githubusercontent.com/MundyMSDS/DATA607/master/inc2edu.csv")

inc <- read_csv(inc, skip = 11)

  kable(head(inc)) %>%
  kable_styling
  
  
```




3.  __Tidy the Data__

#### This step will include renaming, filtering, mutate and gathering activities, oh my!:


```{r echo=TRUE, warning=FALSE, message=FALSE}

inc <- inc %>%
  
  #Rename and select columns
  rename(Bin1 = X1,HS_No = X3,HS_Grad =5, Col_Some = X6, Col_AS=X7,Col_BS =9, Col_MS=10, Col_PD=11, Col_Phd=12) %>% 
  select(-X2) %>% 
  select(-Total) %>% 
  rename(HS_DO =3) %>%
  
  #filter out the non bin rows
  filter(str_detect(Bin1, '\\$')) %>%
  
  #create id number 
  mutate(id = row_number() ) %>%
  
  #create a less granular Bin
  mutate(Bin2 = if_else(id<=10, "Lt24","other")) %>%
  mutate(Bin2 = if_else(id>10 & id <=20 & Bin2=="other", "Lt50",Bin2)) %>%
  mutate(Bin2 = if_else(id>20 & id <=30, "Lt75",Bin2)) %>%
  mutate(Bin2 = if_else(id>30 & id <=40, "Lt99",Bin2)) %>%
  mutate(Bin2 = if_else(id>40, "Mt100",Bin2)) %>% 
  
  #Convert data from wide to skinny
  gather(HS_No, HS_DO, HS_Grad, Col_Some, Col_AS, Col_BS, Col_MS, Col_PD, Col_Phd, key = "Education", value ="N")
    
  kable(head(inc)) %>%
  kable_styling  
  
```

4.  __Explore The Tidy Data Set__

#### I'm going to use a plotly bar chart with a facet_wrap by educational attainment to answer the question.


```{r echo=TRUE, warning=FALSE, message=FALSE}

p <- ggplot(data=inc, aes(x=Bin2, y = N)) +
  geom_bar(stat = "identity") +
  facet_wrap(~ Education) +
  ggtitle("Income vs Educational Attainment - 2018 Census Data")
  

p <- ggplotly(p)
p
  


```



## Answers / Observations

####  Indentify if income and educational attainment are correlated.:

*  The chart above shows a positive correlation between educational attainment and income.
*  High school drop outs and individuals who did not attend high school have the lowest incomes in 2018
*  Highest income go to Bachelor, Masters, Phd and Professal degree holders.
*  There are a relatively small number of Phd and Professional degree holders but salaries are high ($100K) for these degreess.
*  All education levels have some individuals earning more than $100K. 


